### 2023-02-21 219-2023-analysis/manuscript.qmd
use osf storage, update preprocess, add quarto doc

---
title: "Replication of Brydevall et al. 2018"
author: "Josh de Leeuw"
format: html
editor: visual
---

```{r}
#| label: Load libraries
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(ggplot2)
library(readr)
```

## Results

### Behavioral

```{r}
#| label: Load the Behavioral Data
#| echo: false
#| message: false

data.behavioral <- read_csv("data/preprocessed/behavioral.csv")

```

```{r}
#| label: Attention check calculations
#| echo: false

attention.check.responses <- data.behavioral %>%
  filter(task=="catch", phase=="test", card == catch_n) %>%
  group_by(subject_id) %>%
  summarize(successes = sum(rt < 1500), mean_rt = mean(rt))

attention.check.overall.m <- attention.check.responses %>% pull(successes) %>% mean()
attention.check.overall.sd <- attention.check.responses %>% pull(successes) %>%
  sd()
attention.check.rt.m <- attention.check.responses %>% pull(mean_rt) %>%
  mean()
attention.check.rt.sd <- attention.check.responses %>% pull(mean_rt) %>%
  sd()

attention.check.failures <- attention.check.responses %>% filter(successes < 3)
```

There were five attention check trials, one during each block, throughout the experiment. The mean number of correct responses was `r attention.check.overall.m` (*SD* = `r attention.check.overall.sd`) and the average response time was We excluded `r nrow(attention.check.failures)` participants for failing more than 2 attention checks.

### EEG

```{r}
#| label: Load EEG data
#| echo: false
#| message: false

data.eeg <- read_csv('data/preprocessed/eeg.csv')

```

```{r}
#| label: Filter EEG data
#| echo: false

data.eeg.filtered <- data.eeg %>%
  filter(good_segment == TRUE) %>%
  filter(electrode %in% c("Cz", "Fz"))

```

```{r}
#| label: Compute RPEs and IPEs
#| echo: false
#| message: false

hand.information <- data.behavioral %>% 
  filter(phase=='test', task=='reveal') %>%
  mutate(subject = subject_id) %>%
  select(-subject_id) %>%
  group_by(subject) %>%
  mutate(hand_id = rep(1:80, each=5)) %>%
  mutate(losses_so_far = (card-1) - wins_so_far) %>%
  select(subject, hand_id, card, card_value, wins_so_far, losses_so_far, outcome)

hand.information.with.rpe <- hand.information %>%
  mutate(
    wp_before = pbinom(2-losses_so_far, 5-(card-1), 0.5), 
    wp_after = pbinom(2-(losses_so_far+abs(1-card_value)), 5-card, 0.5), 
    rpe = wp_after-wp_before)

entropy <- function(p){
  return(ifelse(p == 0 | p == 1, 0, -p*log2(p) - (1-p)*log2(1-p)))
}

hand.information.with.rpe.and.ipe <- hand.information.with.rpe %>%
  mutate(wp_alternative = pbinom(2-(losses_so_far+card_value), 5-card, 0.5)) %>%
  mutate(information = entropy(wp_before) - entropy(wp_after),
         information_expected = ((entropy(wp_before) - entropy(wp_after)) + (entropy(wp_before) - entropy(wp_alternative))) / 2,
         ipe = information - information_expected)

data.eeg.filtered.rpe.ipe <- data.eeg.filtered %>%
  mutate(card = card_id) %>%
  select(-card_id) %>%
  left_join(hand.information.with.rpe.and.ipe, by=c("subject", "hand_id", "card"))

```

### 2023-02-21 219-2023-analysis/preprocessing/merge-beh.R

@@ -1,9 +1,10 @@
 library(readr)
 library(dplyr)
 library(jsonlite)
+library(tidyr)
 
 all.beh.files <- list.files(path="data/raw/beh", pattern=".json", full.names = T)
 
-all.beh.data <- lapply(all.beh.files, fromJSON) %>% bind_rows()
+all.beh.data <- lapply(all.beh.files, fromJSON) %>% bind_rows() %>% filter(trial_type != "survey-number")
 
 write_csv(all.beh.data, "data/preprocessed/behavioral.csv")

### 2023-02-21 219-2023-analysis/preprocessing/run-preprocess-local.R

@@ -1,12 +1,18 @@
 library(readr)
+library(stringr)
 source("preprocessing/preprocess-eeg-functions.R")
 
-#subjects <- c("01", "02", "03", "04", "05", "08","09","10","11","12","13","14","15")
-#subjects <- c("08","09","10","11","12","13","14","15")
-subjects <- c("14","15")
+subjects <- list.files('data/raw/beh') %>% str_sub(start = 21, end=22)
+
 for(subject in subjects){
-  print(paste0("Working on ",subject))
   
+  final.path <- paste0("data/preprocessed/subject-", subject,"-epochs.csv")
+  print(paste0("Working on ",subject))
+  if(file.exists(final.path)){
+    print(paste0("epochs.csv already exists - skipping"))
+    next
+  }
+  tryCatch({
   eeg.file <- list.files('data/raw/eeg', pattern=paste0("subject-", subject), full.names = T)
   beh.file <- paste0('data/raw/beh/219_2023_behavioral_', subject,'.json')
   
@@ -25,5 +31,11 @@ for(subject in subjects){
     bad_segment_range = 200)
   
   write_csv(preprocessed.data, file=paste0("data/preprocessed/subject-", subject,"-epochs.csv"))
-  
+  },
+  error = function(cond){
+    message(cond)
+  },
+  warning = function(cond){
+    message(cond)
+  })
 }

### 2023-02-21 219-2023-analysis/retrieve-data-from-osf.R

@@ -1,0 +1,12 @@
+library(osfr)
+library(dplyr)
+
+raw_data_component <- osf_retrieve_node("v8tgz")
+
+files <- osf_ls_files(raw_data_component)
+
+eeg.files <- files %>% filter(name=="EEG") %>% osf_ls_files(n_max=Inf)
+beh.files <- files %>% filter(name=="Behavioral") %>% osf_ls_files(n_max=Inf)
+
+osf_download(beh.files, path="data/raw/beh", conflicts = "skip")
+osf_download(eeg.files, path="data/raw/eeg", conflicts = "skip")

### 2023-02-21 219-2023-analysis/test-merged-eeg.R

@@ -5,6 +5,14 @@ library(ggplot2)
 eeg <- read_csv('data/preprocessed/eeg.csv')
 behavioral <- read_csv('data/preprocessed/behavioral.csv')
 
+attention.checks <- behavioral %>% 
+  filter(task=="catch", card==catch_n, phase=="test") %>%
+  select(subject_id, rt, response, task, card, phase)
+
+attention.check.responses <- attention.checks %>%
+  group_by(subject_id) %>%
+  summarize(failed = sum(is.null(rt) | (rt > 1500)))
+
 grand.averages <- eeg %>% filter(electrode %in% c("Cz", "Fz")) %>%
   filter(good_segment==T) %>%
   group_by(subject, t, electrode) %>%

### 2023-02-23 219-2023-analysis/manuscript.qmd
add analysis work

@@
 author: "Josh de Leeuw"
-format: html
+format: 
+  html:
+    df-print: kable
 editor: visual
@@
 library(readr)
+library(ez)
 ```
@@
 attention.check.rt.m <- attention.check.responses %>% pull(mean_rt) %>%
-  mean()
+  mean() %>% round(0)
 attention.check.rt.sd <- attention.check.responses %>% pull(mean_rt) %>%
-  sd()
+  sd() %>% round(0)
 
@@
 There were five attention check trials, one during each block, throughout the experiment.
-The mean number of correct responses was `r attention.check.overall.m` (*SD* = `r attention.check.overall.sd`) and the average response time was We excluded `r nrow(attention.check.failures)` participants for failing more than 2 attention checks.
+The mean number of correct responses was `r attention.check.overall.m` (*SD* = `r attention.check.overall.sd`) and the average response time was `r attention.check.rt.m`ms (*SD* = `r attention.check.rt.sd`).
+As per our pre-registered criteria, we excluded `r nrow(attention.check.failures)` participants for failing more than 2 attention checks (defined as a response time greater than 1,500ms).
 
@@
          information_expected = ((entropy(wp_before) - entropy(wp_after)) + (entropy(wp_before) - entropy(wp_alternative))) / 2,
-         ipe = information - information_expected)
+         ipe = information - information_expected) %>%
+  mutate(rpe_type = case_when(
+    rpe > 0 ~ "Positive",
+    rpe < 0 ~ "Negative",
+    rpe == 0 ~ "None")) %>%
+  mutate(ipe_type = case_when(
+    ipe > 0 ~ "Positive",
+    ipe < 0 ~ "Negative",
+    ipe == 0 ~ "None"))
 
@@
 ```
+
+```{r}
+#| label: Define ERP time window
+#| echo: false
+
+frn.window <- 200:350
+```
+
+```{r}
+#| label: Count usable segments for RPEs
+#| echo: false
+#| message: false
+
+rpe.segments <- data.eeg.filtered.rpe.ipe %>%
+  filter(rpe_type != "None") %>%
+  group_by(subject, electrode, hand_id, card, rpe_type) %>%
+  summarize(n = n()) %>%
+  group_by(subject, electrode, rpe_type) %>%
+  summarize(n = n())
+```
+
+```{r}
+#| label: Count usable segments for IPEs
+#| echo: false
+#| message: false
+
+ipe.segments <- data.eeg.filtered.rpe.ipe %>%
+  filter(ipe_type != "None") %>%
+  group_by(subject, electrode, hand_id, card, ipe_type) %>%
+  summarize(n = n()) %>%
+  group_by(subject, electrode, ipe_type) %>%
+  summarize(n = n())
+```
+
+We pre-registered an inclusion criteria that a participant must have at least 20 usable epochs of data of a particular event type in order to be included in the corresponding analyses.
+The minimum number of usable epochs for the RPE analysis in an individual subject was `r rpe.segments %>% pull(n) %>% min()`, *M*=`r rpe.segments %>% pull(n) %>% mean() %>% round(0)`, *SD*=`r rpe.segments %>% pull(n) %>% sd() %>% round(0)`.
+The minimum number of usable epochs for the IPE analysis in an individual subject was `r ipe.segments %>% pull(n) %>% min()`, *M*=`r ipe.segments %>% pull(n) %>% mean() %>% round(0)`, *SD*=`r ipe.segments %>% pull(n) %>% sd() %>% round(0)`.
+Therefore, no participants needed to be excluded from the following analyses.
+
+```{r}
+#| label: Generate data for Grand Averages
+#| echo: false
+#| message: false
+
+rpe.erp.data.subject <- data.eeg.filtered.rpe.ipe %>%
+  filter(rpe_type != "None") %>%
+  group_by(subject, electrode, t, rpe_type) %>%
+  summarize(m.subject = mean(v))
+
+rpe.erp.data.group <- rpe.erp.data.subject %>%
+  group_by(electrode, t, rpe_type) %>%
+  summarize(m = mean(m.subject), se = sd(m.subject)/sqrt(n())) %>%
+  mutate(error_type = "Reward Prediction Error", error_sign = rpe_type)
+
+ipe.erp.data.subject <- data.eeg.filtered.rpe.ipe %>%
+  filter(ipe_type != "None") %>%
+  group_by(subject, electrode, t, ipe_type) %>%
+  summarize(m.subject = mean(v))
+
+ipe.erp.data.group <- ipe.erp.data.subject %>%
+  group_by(electrode, t, ipe_type) %>%
+  summarize(m = mean(m.subject), se = sd(m.subject)/sqrt(n())) %>%
+  mutate(error_type = "Information Prediction Error", error_sign = ipe_type)
+
+erp.data <- rpe.erp.data.group %>% bind_rows(ipe.erp.data.group) %>%
+  select(-rpe_type, -ipe_type)
+
+```
+
+```{r}
+#| label: fig-grand-averages
+#| echo: false
+#| fig-cap: Grand average ERPs for both electrode sites (Cz, Fz) and both error types (Information Prediction and Reward Prediction).
+The critical time window for our analysis, 200-350ms, is highlighted.
+
+ggplot(erp.data, aes(x=t, y=m, ymin=m-se, ymax=m+se, color=error_sign, fill=error_sign))+
+  facet_grid(error_type ~ electrode)+
+  geom_hline(yintercept = 0)+
+  geom_vline(xintercept = 0)+
+  annotate("rect", xmin=min(frn.window), xmax=max(frn.window), ymin=-Inf, ymax=Inf, alpha=0.4)+
+  geom_ribbon(color=NA, alpha = 0.2)+
+  geom_line()+
+  scale_y_reverse()+
+  scale_color_brewer(type="qual", palette = "Set1", name="Error Direction")+
+  scale_fill_brewer(type="qual", palette = "Set1", name="Error Direction")+
+  labs(x="Time since card value revealed (ms)", y="Microvolts")+
+  theme_minimal()+
+  theme(panel.grid = element_blank())
+
+```
+
+```{r}
+#| label: Compute ANOVA data
+#| echo: false
+#| message: false
+
+rpe.anova.data <- rpe.erp.data.subject %>%
+  filter(t %in% frn.window) %>%
+  group_by(subject, electrode, rpe_type) %>%
+  summarize(M = mean(m.subject))
+
+ipe.anova.data <- ipe.erp.data.subject %>%
+  filter(t %in% frn.window) %>%
+  group_by(subject, electrode, ipe_type) %>%
+  summarize(M = mean(m.subject))
+```
+
+```{r}
+#| label: RPE ANOVA
+#| echo: false
+#| warning: false
+
+rpe.anova <- ezANOVA(rpe.anova.data, dv=M, wid=subject, within=c(electrode, rpe_type))
+
+rpe.anova$ANOVA
+```
+
+```{r}
+#| label: IPE ANOVA
+#| echo: false
+#| warning: false
+
+ipe.anova <- ezANOVA(ipe.anova.data, dv=M, wid=subject, within=c(electrode, ipe_type))
+
+ipe.anova$ANOVA
+```

### 2023-02-23 219-2023-analysis/preprocessing/merge-beh.R

@@ -5,6 +5,6 @@ library(tidyr)
 
 all.beh.files <- list.files(path="data/raw/beh", pattern=".json", full.names = T)
 
-all.beh.data <- lapply(all.beh.files, fromJSON) %>% bind_rows() %>% filter(trial_type != "survey-number")
+all.beh.data <- lapply(all.beh.files, fromJSON) %>% bind_rows() %>% dplyr::filter(trial_type != "survey-number")
 
 write_csv(all.beh.data, "data/preprocessed/behavioral.csv")

### 2023-02-23 219-2023-analysis/preprocessing/merge-eeg.R

@@ -1,7 +1,7 @@
 library(readr)
 library(dplyr)
 
-all.eeg.epoch.files <- list.files(path="data/preprocessed", pattern=".csv", full.names = T)
+all.eeg.epoch.files <- list.files(path="data/preprocessed", pattern="epochs.csv", full.names = T)
 
 all.eeg.epochs <- lapply(all.eeg.epoch.files, read_csv, col_types="iiicdddllc")
 

### 2023-02-23 219-2023-analysis/preprocessing/preprocess-eeg-functions.R

@@ -1,11 +1,11 @@
-library(eegkit)
 library(dplyr)
 library(tidyr)
 library(purrr)
 library(ravetools)
+library(eegkit)
 source('preprocessing/extract-events.R')
 
-preprocess_eeg <- function(file, beh.file, subject_id, sampling_rate=500, filter_low=0.1, filter_high=70, notch_filter_low=58, notch_filter_high=62, segment_begin=-100, segment_end=1000, segment_offset=0, bad_segment_range=200, which_electrodes=c('Cz', 'Fz', 'Fp1', 'Fp2')){
+preprocess_eeg <- function(file, beh.file, subject_id, sampling_rate=500, filter_low=0.1, filter_high=70, notch_filter_low=59, notch_filter_high=61, segment_begin=-100, segment_end=1000, segment_offset=0, bad_segment_range=200, which_electrodes=c('Cz', 'Fz', 'Fp1', 'Fp2')){
   data <- read_eeg_tidy(eeg.file, beh.file, which_electrodes) %>%
     linked_ears_rereference() %>%
     bandpass(low=filter_low, high=filter_high, sampling_rate=sampling_rate) %>%
@@ -57,7 +57,7 @@ bandpass <- function(data, low, high, sampling_rate){
   df <- data$signals
   df <- df %>% 
     group_by(electrode) %>% 
-    mutate(v = band_pass1(v, sampling_rate, low, high)) %>%
+    mutate(v = eegfilter(v, sampling_rate, low, high)) %>%
     ungroup()
   data$signals <- df
   return(data)

### 2023-02-24 219-2023-analysis/manuscript.qmd
use osf to get data

@@
 
-## Results
-
-### Behavioral
+## Methods
 
@@
 data.behavioral <- read_csv("data/preprocessed/behavioral.csv")
+```
 
+### Participants
+
+```{r}
+#| label: Count participants
+#| echo: false
+
+n.participants <- data.behavioral %>% pull(subject_id) %>% unique() %>% length()
 ```
+
+`r n.participants` Vassar College students participated in this study. All participants were right-handed.
+
+## Results
+
+### Behavioral
 
@@
 
+```{r}
+#| label: Exclude bad EEG
+#| echo: false
+
+bad.eeg.subjects <- rpe.segments %>% 
+  filter(n < 20) %>% 
+  bind_rows(ipe.segments %>% filter(n < 20)) %>%
+  pull(subject) %>%
+  unique()
+```
+
 We pre-registered an inclusion criteria that a participant must have at least 20 usable epochs of data of a particular event type in order to be included in the corresponding analyses.
@@
 The minimum number of usable epochs for the IPE analysis in an individual subject was `r ipe.segments %>% pull(n) %>% min()`, *M*=`r ipe.segments %>% pull(n) %>% mean() %>% round(0)`, *SD*=`r ipe.segments %>% pull(n) %>% sd() %>% round(0)`.
-Therefore, no participants needed to be excluded from the following analyses.
+`r bad.eeg.subjects %>% length()` subject was excluded for poor EEG data quality.
 
@@
   filter(rpe_type != "None") %>%
+  filter(!subject %in% bad.eeg.subjects) %>%
   group_by(subject, electrode, t, rpe_type) %>%
@@
   filter(ipe_type != "None") %>%
+  filter(!subject %in% bad.eeg.subjects) %>%
   group_by(subject, electrode, t, ipe_type) %>%

### 2023-02-28 219-2023-analysis/test-alignment.R
add alignment test file

@@ -1,0 +1,30 @@
+library(text.alignment)
+
+a <- events.hands.only$TRIGGER
+b <- hand.ids$sequence_id
+
+a <- a[c(1:15,17:81)]
+
+L <- c(LETTERS, letters)
+
+aa <- L[a]
+bb <- L[b]
+
+aaa <- paste0(aa, collapse = "")
+bbb <- paste0(bb, collapse = "")
+
+zz <- text.alignment::smith_waterman(aaa,bbb,lower=FALSE)
+
+
+# aaa has the extra event, need to figure out how to discard it
+
+zz$a
+
+a.z <- unlist(sapply(stringr::str_split(zz$a$alignment$text,"")[[1]], function(x){ 
+  if(x=="#") { return(NA) } 
+  return(which(L==x))
+  }, simplify=TRUE))
+b.z <- unlist(sapply(stringr::str_split(zz$b$alignment$text,"")[[1]], function(x){ 
+  if(x=="#") { return(NA) } 
+  return(which(L==x))
+  }, simplify=TRUE))

### 2023-02-28 219-2023-analysis/manuscript.qmd
use osf to get data
add osfr as dependency

@@
 library(ez)
+library(osfr)
+```
+
+```{r}
+#| label: Get preprocessed data from OSF
+#| include: false
+#| echo: false
+
+osf_retrieve_node("2pxf4") %>%
+  osf_ls_files() %>%
+  osf_download(path="data/preprocessed/", conflicts="skip")
 ```

### 2023-02-28 219-2023-analysis/preprocessing/extract-events.R
implement better alignment algorithm

@@ -40,14 +40,57 @@ extract_events <- function(eeg.file, beh.file){
   hand_counter <- 1
   eeg_event_counter <- 1
   
-  while(eeg_event_counter <= nrow(events.hands.only)){
-    if(events.hands.only$TRIGGER[eeg_event_counter] == hand.ids$sequence_id[hand_counter]){
-      events.hands.only$hand_id[eeg_event_counter] = hand.ids$hand_id[hand_counter]
-      eeg_event_counter <- eeg_event_counter + 1
+  # while(eeg_event_counter <= nrow(events.hands.only)){
+  #   if(events.hands.only$TRIGGER[eeg_event_counter] == hand.ids$sequence_id[hand_counter]){
+  #     events.hands.only$hand_id[eeg_event_counter] = hand.ids$hand_id[hand_counter]
+  #     eeg_event_counter <- eeg_event_counter + 1
+  #   } 
+  #   hand_counter <- hand_counter + 1
+  # }
+  
+  eeg.hand.events <- events.hands.only$TRIGGER
+  jspsych.hand.sequence <- hand.ids$sequence_id
+  
+  L <- c(LETTERS, letters)
+  
+  eeg.hand.events.encoded  <- L[eeg.hand.events] %>% paste0(collapse="")
+  jspsych.hand.sequence.encoded <- L[jspsych.hand.sequence] %>% paste0(collapse="")
+  
+  alignment.solution <- text.alignment::smith_waterman(eeg.hand.events.encoded, jspsych.hand.sequence.encoded, lower=FALSE)
+  
+  a.z <- unlist(sapply(stringr::str_split(alignment.solution$a$alignment$text,"")[[1]], function(x){ 
+    if(x=="#") { return(NA) } 
+    return(which(L==x))
+  }, simplify=TRUE))
+  b.z <- unlist(sapply(stringr::str_split(alignment.solution$b$alignment$text,"")[[1]], function(x){ 
+    if(x=="#") { return(NA) } 
+    return(which(L==x))
+  }, simplify=TRUE))
+  
+  aligned.hands <- rep(0, length(a))
+  event.loc <- 1
+  hand.id <- 1
+  for(i in 1:length(a.z)){
+    if(is.na(b.z[i])) { 
+      aligned.hands[event.loc] <- NA 
+      event.loc <- event.loc + 1
+      next
     } 
-    hand_counter <- hand_counter + 1
+    if(is.na(a.z[i])) { 
[…]

### 2023-02-28 219-2023-analysis/test-alignment.R

@@ -1,30 +1,42 @@
 library(text.alignment)
 
-a <- events.hands.only$TRIGGER
-b <- hand.ids$sequence_id
-
-a <- a[c(1:15,17:81)]
+eeg.hand.events <- events.hands.only$TRIGGER
+jspsych.hand.sequence <- hand.ids$sequence_id
 
 L <- c(LETTERS, letters)
 
-aa <- L[a]
-bb <- L[b]
+eeg.hand.events.encoded  <- L[eeg.hand.events] %>% paste0(collapse="")
+jspsych.hand.sequence.encoded <- L[jspsych.hand.sequence] %>% paste0(collapse="")
 
-aaa <- paste0(aa, collapse = "")
-bbb <- paste0(bb, collapse = "")
+alignment.solution <- text.alignment::smith_waterman(eeg.hand.events.encoded, jspsych.hand.sequence.encoded, lower=FALSE)
 
-zz <- text.alignment::smith_waterman(aaa,bbb,lower=FALSE)
-
-
-# aaa has the extra event, need to figure out how to discard it
-
-zz$a
-
-a.z <- unlist(sapply(stringr::str_split(zz$a$alignment$text,"")[[1]], function(x){ 
+a.z <- unlist(sapply(stringr::str_split(alignment.solution$a$alignment$text,"")[[1]], function(x){ 
   if(x=="#") { return(NA) } 
   return(which(L==x))
   }, simplify=TRUE))
-b.z <- unlist(sapply(stringr::str_split(zz$b$alignment$text,"")[[1]], function(x){ 
+b.z <- unlist(sapply(stringr::str_split(alignment.solution$b$alignment$text,"")[[1]], function(x){ 
   if(x=="#") { return(NA) } 
   return(which(L==x))
   }, simplify=TRUE))
+
+aligned.hands <- rep(0, length(a))
+event.loc <- 1
+hand.id <- 1
+for(i in 1:length(a.z)){
+  if(is.na(b.z[i])) { 
+    aligned.hands[event.loc] <- NA 
+    event.loc <- event.loc + 1
+    next
+  } 
+  if(is.na(a.z[i])) { 
+    hand.id <- hand.id + 1
+    next
+  }
+  if(a.z[i] == b.z[i]) { 
+    aligned.hands[event.loc] <- hand.id 
+    event.loc <- event.loc + 1
+    hand.id <- hand.id + 1
+    next
+  }
+}
+  

### 2023-03-02 219-2023-analysis/manuscript.qmd
updated preprocessing, ms

@@
 
-n.participants <- data.behavioral %>% pull(subject_id) %>% unique() %>% length()
+n.participants <- data.behavioral %>% pull(subject) %>% unique() %>% length()
 ```
@@
 attention.check.responses <- data.behavioral %>%
-  filter(task=="catch", phase=="test", card == catch_n) %>%
-  group_by(subject_id) %>%
-  summarize(successes = sum(rt < 1500), mean_rt = mean(rt))
+  filter(task=="catch", phase=="test", card_id == catch_n) %>%
+  group_by(subject) %>%
+  summarize(successes = sum(rt < 1500, na.rm=TRUE), mean_rt = mean(rt, na.rm=TRUE))
 
-attention.check.overall.m <- attention.check.responses %>% pull(successes) %>% mean()
-attention.check.overall.sd <- attention.check.responses %>% pull(successes) %>%
-  sd()
+attention.check.tally <- attention.check.responses %>% group_by(successes) %>% summarize(n=n())
+
 attention.check.rt.m <- attention.check.responses %>% pull(mean_rt) %>%
@@
 There were five attention check trials, one during each block, throughout the experiment.
-The mean number of correct responses was `r attention.check.overall.m` (*SD* = `r attention.check.overall.sd`) and the average response time was `r attention.check.rt.m`ms (*SD* = `r attention.check.rt.sd`).
+`r attention.check.tally %>% filter(successes==5) %>% pull(n)` of `r n.participants` responded in time to all 5 attention checks, and `r attention.check.tally %>% filter(successes==4) %>% pull(n)` responded correctly to 4 of the 5 attention checks.
+The average response time was `r attention.check.rt.m`ms (*SD* = `r attention.check.rt.sd`).
 As per our pre-registered criteria, we excluded `r nrow(attention.check.failures)` participants for failing more than 2 attention checks (defined as a response time greater than 1,500ms).
+
+```{r}
+#| label: Compute RPEs and IPEs
+#| echo: false
+#| message: false
+
+hand.information <- data.behavioral %>% 
+  filter(phase=='test', task=='reveal') %>%
+  group_by(subject) %>%
+  mutate(hand_id = rep(1:80, each=5)) %>%
+  ungroup() %>%
+  mutate(losses_so_far = (card_id-1) - wins_so_far) %>%
+  select(subject, hand_id, card_id, card_value, wins_so_far, losses_so_far, outcome)
+
+hand.information.with.rpe <- hand.information %>%
+  mutate(
+    wp_before = pbinom(2-losses_so_far, 5-(card_id-1), 0.5), 
+    wp_after = pbinom(2-(losses_so_far+(1-card_value)), 5-card_id, 0.5), 
+    rpe = wp_after-wp_before)
+
+entropy <- function(p){
+  return(ifelse(p == 0 | p == 1, 0, -p*log2(p) - (1-p)*log2(1-p)))
+}
+
+hand.information.with.rpe.and.ipe <- hand.information.with.rpe %>%
+  mutate(wp_alternative = pbinom(2-(losses_so_far+card_value), 5-card_id, 0.5)) %>%
+  mutate(information = entropy(wp_before) - entropy(wp_after),
+         information_alt = entropy(wp_before) - entropy(wp_alternative),
+         information_expected = (information + information_alt) / 2,
+         ipe = information - information_expected) %>%
+  mutate(rpe_type = case_when(
+    rpe > 0 ~ "Positive",
+    rpe < 0 ~ "Negative",
+    rpe == 0 ~ "None")) %>%
+  mutate(ipe_type = case_when(
+    ipe > 0 ~ "Positive",
+    ipe < 0 ~ "Negative",
+    ipe == 0 ~ "None"))
+```
 
@@
 ```{r}
-#| label: Compute RPEs and IPEs
+#| label: Merge RPE and IPE info into EEG data
 #| echo: false
@@
 
-hand.information <- data.behavioral %>% 
-  filter(phase=='test', task=='reveal') %>%
-  mutate(subject = subject_id) %>%
-  select(-subject_id) %>%
-  group_by(subject) %>%
-  mutate(hand_id = rep(1:80, each=5)) %>%
-  mutate(losses_so_far = (card-1) - wins_so_far) %>%
-  select(subject, hand_id, card, card_value, wins_so_far, losses_so_far, outcome)
+data.eeg.filtered.rpe.ipe <- data.eeg.filtered %>%
+  left_join(hand.information.with.rpe.and.ipe, by=c("subject", "hand_id", "card_id"))
 
-hand.information.with.rpe <- hand.information %>%
-  mutate(
-    wp_before = pbinom(2-losses_so_far, 5-(card-1), 0.5), 
-    wp_after = pbinom(2-(losses_so_far+abs(1-card_value)), 5-card, 0.5), 
-    rpe = wp_after-wp_before)
-
-entropy <- function(p){
-  return(ifelse(p == 0 | p == 1, 0, -p*log2(p) - (1-p)*log2(1-p)))
-}
-
-hand.information.with.rpe.and.ipe <- hand.information.with.rpe %>%
-  mutate(wp_alternative = pbinom(2-(losses_so_far+card_value), 5-card, 0.5)) %>%
-  mutate(information = entropy(wp_before) - entropy(wp_after),
-         information_expected = ((entropy(wp_before) - entropy(wp_after)) + (entropy(wp_before) - entropy(wp_alternative))) / 2,
-         ipe = information - information_expected) %>%
-  mutate(rpe_type = case_when(
-    rpe > 0 ~ "Positive",
-    rpe < 0 ~ "Negative",
-    rpe == 0 ~ "None")) %>%
-  mutate(ipe_type = case_when(
-    ipe > 0 ~ "Positive",
-    ipe < 0 ~ "Negative",
-    ipe == 0 ~ "None"))
-
-data.eeg.filtered.rpe.ipe <- data.eeg.filtered %>%
-  mutate(card = card_id) %>%
-  select(-card_id) %>%
-  left_join(hand.information.with.rpe.and.ipe, by=c("subject", "hand_id", "card"))
-
-```
-
-```{r}
-#| label: Define ERP time window
-#| echo: false
-
-frn.window <- 200:350
 ```
@@
   filter(rpe_type != "None") %>%
-  group_by(subject, electrode, hand_id, card, rpe_type) %>%
-  summarize(n = n()) %>%
+  group_by(subject, electrode, hand_id, card_id, rpe_type) %>%
+  slice_head(n=1) %>%
   group_by(subject, electrode, rpe_type) %>%
-  summarize(n = n())
+  summarize(n=n())
 ```
@@
   filter(ipe_type != "None") %>%
-  group_by(subject, electrode, hand_id, card, ipe_type) %>%
-  summarize(n = n()) %>%
+  group_by(subject, electrode, hand_id, card_id, ipe_type) %>%
+  slice_head(n=1) %>%
   group_by(subject, electrode, ipe_type) %>%
@@
 We pre-registered an inclusion criteria that a participant must have at least 20 usable epochs of data of a particular event type in order to be included in the corresponding analyses.
-The minimum number of usable epochs for the RPE analysis in an individual subject was `r rpe.segments %>% pull(n) %>% min()`, *M*=`r rpe.segments %>% pull(n) %>% mean() %>% round(0)`, *SD*=`r rpe.segments %>% pull(n) %>% sd() %>% round(0)`.
-The minimum number of usable epochs for the IPE analysis in an individual subject was `r ipe.segments %>% pull(n) %>% min()`, *M*=`r ipe.segments %>% pull(n) %>% mean() %>% round(0)`, *SD*=`r ipe.segments %>% pull(n) %>% sd() %>% round(0)`.
-`r bad.eeg.subjects %>% length()` subject was excluded for poor EEG data quality.
+`r bad.eeg.subjects %>% length()` subjects were excluded for having fewer than 20 usable epochs in all event types.
+Of the remaining subjects, the average number of usable epochs for the RPE analysis in an individual subject was `r rpe.segments %>% filter(!subject %in% bad.eeg.subjects) %>% pull(n) %>% mean() %>% round(0)`, *SD*=`r rpe.segments %>% filter(!subject %in% bad.eeg.subjects) %>% pull(n) %>% sd() %>% round(0)`.
+The average number of usable epochs for the IPE analysis was `r ipe.segments %>% filter(!subject %in% bad.eeg.subjects) %>% pull(n) %>% mean() %>% round(0)`, *SD*=`r ipe.segments %>% filter(!subject %in% bad.eeg.subjects) %>% pull(n) %>% sd() %>% round(0)`.
+
+```{r}
+#| label: Define ERP time window
+#| echo: false
+
+frn.window <- 200:350
+```
 

### 2023-03-02 219-2023-analysis/behavioral-data-test.R

@@ -1,8 +1,0 @@
-library(jsonlite)
-library(dplyr)
-
-df <- fromJSON('data/behavior/219_2023_behavioral_01.json')
-
-card_reveals <- df %>% filter(task=="reveal") %>% select(eeg_event_id, card, color, participant_choice, card_value, wins_so_far, outcome, sequence_id)
-
-beh_hands <- card_reveals[seq(11,410,5),]

### 2023-03-02 219-2023-analysis/preprocessing/extract-events.R

@@ -37,9 +37,6 @@ extract_events <- function(eeg.file, beh.file){
     mutate(hand_id = NA) %>%
     select(-time)
   
-  hand_counter <- 1
-  eeg_event_counter <- 1
-  
   # while(eeg_event_counter <= nrow(events.hands.only)){
   #   if(events.hands.only$TRIGGER[eeg_event_counter] == hand.ids$sequence_id[hand_counter]){
   #     events.hands.only$hand_id[eeg_event_counter] = hand.ids$hand_id[hand_counter]
@@ -56,18 +53,19 @@ extract_events <- function(eeg.file, beh.file){
   eeg.hand.events.encoded  <- L[eeg.hand.events] %>% paste0(collapse="")
   jspsych.hand.sequence.encoded <- L[jspsych.hand.sequence] %>% paste0(collapse="")
   
-  alignment.solution <- text.alignment::smith_waterman(eeg.hand.events.encoded, jspsych.hand.sequence.encoded, lower=FALSE)
+  #alignment.solution <- text.alignment::smith_waterman(eeg.hand.events.encoded, jspsych.hand.sequence.encoded, gap=-1,lower=FALSE)
+  alignment.solution <- NameNeedle::needles(eeg.hand.events.encoded, jspsych.hand.sequence.encoded)
   
-  a.z <- unlist(sapply(stringr::str_split(alignment.solution$a$alignment$text,"")[[1]], function(x){ 
-    if(x=="#") { return(NA) } 
+  a.z <- unlist(sapply(stringr::str_split(alignment.solution$align1,"")[[1]], function(x){ 
+    if(x=="*") { return(NA) } 
     return(which(L==x))
   }, simplify=TRUE))
-  b.z <- unlist(sapply(stringr::str_split(alignment.solution$b$alignment$text,"")[[1]], function(x){ 
-    if(x=="#") { return(NA) } 
+  b.z <- unlist(sapply(stringr::str_split(alignment.solution$align2,"")[[1]], function(x){ 
+    if(x=="*") { return(NA) } 
     return(which(L==x))
   }, simplify=TRUE))
   
-  aligned.hands <- rep(0, length(a))
+  aligned.hands <- rep(0, length(eeg.hand.events.encoded))
   event.loc <- 1
   hand.id <- 1
   for(i in 1:length(a.z)){
@@ -103,9 +101,17 @@ extract_events <- function(eeg.file, beh.file){
     dplyr::filter(!is.na(offset_time)) %>%
     rowwise() %>%
[…]

### 2023-03-02 219-2023-analysis/preprocessing/merge-beh.R

@@ -5,6 +5,12 @@ library(tidyr)
 
 all.beh.files <- list.files(path="data/raw/beh", pattern=".json", full.names = T)
 
-all.beh.data <- lapply(all.beh.files, fromJSON) %>% bind_rows() %>% dplyr::filter(trial_type != "survey-number")
+all.beh.data <- lapply(all.beh.files, fromJSON) %>% bind_rows() %>% 
+  dplyr::filter(trial_type != "survey-number") %>%
+  rename(subject=subject_id, card_id=card) %>%
+  mutate(outcome = case_when(
+    outcome == 0 ~ "win",
+    outcome == 1 ~ "loss",
+    .default = NA))
 
 write_csv(all.beh.data, "data/preprocessed/behavioral.csv")

### 2023-03-02 219-2023-analysis/preprocessing/run-eeg-preprocess.R

@@ -1,25 +1,0 @@
-SUBJECT <- "03"
-eeg.file <- list.files('data/eeg', pattern=paste0("subject-", SUBJECT), full.names = T)
-beh.file <- paste0('data/behavior/219_2023_behavioral_', SUBJECT,'.json')
-
-source('preprocessing/preprocess-eeg-functions.R')
-
-result <- preprocess_eeg(eeg.file, beh.file, SUBJECT)
-
-
-library(ggplot2)
-
-segment.eeg <- result %>% filter(hand_id == 50, card_id == 4, electrode == 'Fp1')
-
-ggplot(segment.eeg, aes(x=t, y=v)) +
-  geom_hline(yintercept=0)+
-  geom_vline(xintercept=0)+
-  geom_line() +
-  theme_bw()
-
-blinks <- result %>% filter(eyeblink == TRUE, electrode == 'Fp1')
-
-ggplot(blinks, aes(x=t, y=v))+
-  facet_wrap(vars(hand_id, card_id))+
-  geom_line()+
-  theme_bw()

### 2023-03-02 219-2023-analysis/preprocessing/run-hpc-eeg.sh

@@ -1,28 +1,0 @@
-#!/bin/bash
-#SBATCH --job-name=219eeg
-#SBATCH --output=log-%j.txt
-#SBATCH --nodes=1
-#SBATCH --ntasks=15
-#SBACTH --ntasks-per-node=15
-#SBATCH --partition=general
-#SBATCH --time=99:00:00
-#SBATCH --mail-type=END  	  # Type of email notification- BEGIN,END,FAIL,ALL
-#SBATCH --mail-user=jdeleeuw@vassar.edu
-
-cd ~/219-2023
-
-srun --ntasks=1 Rscript run-preprocess-hpc.R 01 &
-srun --ntasks=1 Rscript run-preprocess-hpc.R 02 &
-srun --ntasks=1 Rscript run-preprocess-hpc.R 03 &
-srun --ntasks=1 Rscript run-preprocess-hpc.R 04 &
-srun --ntasks=1 Rscript run-preprocess-hpc.R 05 &
-srun --ntasks=1 Rscript run-preprocess-hpc.R 08 &
-srun --ntasks=1 Rscript run-preprocess-hpc.R 09 &
-srun --ntasks=1 Rscript run-preprocess-hpc.R 10 &
-srun --ntasks=1 Rscript run-preprocess-hpc.R 11 &
-srun --ntasks=1 Rscript run-preprocess-hpc.R 12 &
-srun --ntasks=1 Rscript run-preprocess-hpc.R 13 &
-srun --ntasks=1 Rscript run-preprocess-hpc.R 14 &
-srun --ntasks=1 Rscript run-preprocess-hpc.R 15 &
-
-wait

### 2023-03-02 219-2023-analysis/preprocessing/run-preprocess-hpc.R

@@ -1,27 +1,0 @@
-library(readr)
-
-source("preprocess-eeg-functions.R")
-
-args <- commandArgs(trailingOnly = TRUE)
-subject <- args[1]
-
-print(paste0("Working on ",subject))
-
-eeg.file <- list.files('/scratch/jdeleeuw/219-2023-eeg/raw/eeg', pattern=paste0("subject-", SUBJECT), full.names = T)
-beh.file <- paste0('/scratch/jdeleeuw/219-2023-eeg/raw/beh/219_2023_behavioral_', SUBJECT,'.json')
-
-preprocessed.data <- preprocess_eeg(
-  file = eeg.file, 
-  beh.file = beh.file,
-  subject_id = subject,
-  sampling_rate = 500,
-  filter_low = 0.1,
-  filter_high = 70,
-  notch_filter_low = 59,
-  notch_filter_high = 61,
-  segment_begin = -100,
-  segment_end = 1000,
-  segment_offset = 0,
-  bad_segment_range = 200)
-
-write_csv(preprocessed.data, file=paste0("/scratch/jdeleeuw/219-2023-eeg/preprocessed/subject-", subject,"-epochs.csv"))

### 2023-03-02 219-2023-analysis/retrieve-data-from-osf.R

@@ -5,8 +5,8 @@ raw_data_component <- osf_retrieve_node("v8tgz")
 
 files <- osf_ls_files(raw_data_component)
 
-eeg.files <- files %>% filter(name=="EEG") %>% osf_ls_files(n_max=Inf)
-beh.files <- files %>% filter(name=="Behavioral") %>% osf_ls_files(n_max=Inf)
+eeg.files <- files %>% dplyr::filter(name=="EEG") %>% osf_ls_files(n_max=Inf)
+beh.files <- files %>% dplyr::filter(name=="Behavioral") %>% osf_ls_files(n_max=Inf)
 
 osf_download(beh.files, path="data/raw/beh", conflicts = "skip")
 osf_download(eeg.files, path="data/raw/eeg", conflicts = "skip")

### 2023-03-02 219-2023-analysis/test-alignment.R

@@ -1,42 +1,0 @@
-library(text.alignment)
-
-eeg.hand.events <- events.hands.only$TRIGGER
-jspsych.hand.sequence <- hand.ids$sequence_id
-
-L <- c(LETTERS, letters)
-
-eeg.hand.events.encoded  <- L[eeg.hand.events] %>% paste0(collapse="")
-jspsych.hand.sequence.encoded <- L[jspsych.hand.sequence] %>% paste0(collapse="")
-
-alignment.solution <- text.alignment::smith_waterman(eeg.hand.events.encoded, jspsych.hand.sequence.encoded, lower=FALSE)
-
-a.z <- unlist(sapply(stringr::str_split(alignment.solution$a$alignment$text,"")[[1]], function(x){ 
-  if(x=="#") { return(NA) } 
-  return(which(L==x))
-  }, simplify=TRUE))
-b.z <- unlist(sapply(stringr::str_split(alignment.solution$b$alignment$text,"")[[1]], function(x){ 
-  if(x=="#") { return(NA) } 
-  return(which(L==x))
-  }, simplify=TRUE))
-
-aligned.hands <- rep(0, length(a))
-event.loc <- 1
-hand.id <- 1
-for(i in 1:length(a.z)){
-  if(is.na(b.z[i])) { 
-    aligned.hands[event.loc] <- NA 
-    event.loc <- event.loc + 1
-    next
-  } 
-  if(is.na(a.z[i])) { 
-    hand.id <- hand.id + 1
-    next
-  }
-  if(a.z[i] == b.z[i]) { 
-    aligned.hands[event.loc] <- hand.id 
-    event.loc <- event.loc + 1
-    hand.id <- hand.id + 1
-    next
-  }
-}
-  

### 2023-03-02 219-2023-analysis/test-event-extraction.R

@@ -1,68 +1,0 @@
-library(edfReader)
-library(purrr)
-library(dplyr)
-library(jsonlite)
-
-
-# subject?
-
-SUBJECT <- "03"
-
-# read in EEG data
-
-eeg.file <- list.files('data/eeg', pattern=paste0("subject-", SUBJECT), full.names = T)
-head <- readEdfHeader(eeg.file)
-signals <- readEdfSignals(head, signals="Ordinary")
-
-
-eeg.data <- map_df(signals, "signal")
-eeg.data$sample_id <- 1:nrow(eeg.data)
-
-## read in behavioral data
-
-behavioral.data <- fromJSON(paste0('data/behavior/219_2023_behavioral_', SUBJECT,'.json'))
-
-card.reveals <- behavioral.data %>% filter(task=="reveal") %>% select(eeg_event_id, card, color, participant_choice, card_value, wins_so_far, outcome, sequence_id)
-
-hand.ids <- card.reveals[seq(11,410,5),] %>%
-  mutate(hand_id = 1:n())
-
-## extract EEG events
-
-events <- eeg.data %>% 
-  select(sample_id, TRIGGER) %>% 
-  filter(TRIGGER != lag(TRIGGER)) %>%
-  filter(TRIGGER > 0) %>%
-  mutate(time = sample_id/500)
-
-events.hands.only <- events %>% 
-  filter(TRIGGER <= 32) %>% 
-  mutate(hand_start_time = time) %>%
-  mutate(hand_end_time = time + 17) %>%
-  mutate(hand_id = NA) %>%
-  select(-time)
-
-hand_counter <- 1
-eeg_event_counter <- 1
-
-while(eeg_event_counter <= nrow(events.hands.only)){
-  if(events.hands.only$TRIGGER[eeg_event_counter] == hand.ids$sequence_id[hand_counter]){
-    events.hands.only$hand_id[eeg_event_counter] = hand.ids$hand_id[hand_counter]
-    eeg_event_counter <- eeg_event_counter + 1
-  } 
-  hand_counter <- hand_counter + 1
-}
-
-
-events.flips.only <- events %>% 
-  filter(TRIGGER == 65535)
-
-events.flips.with.hand.id <- events.flips.only %>%
-  left_join(events.hands.only %>% select(hand_start_time, hand_end_time, hand_id), by=join_by(between(x$time, y$hand_start_time, y$hand_end_time)))
-
[…]

### 2023-03-02 219-2023-analysis/test-merged-eeg.R

@@ -1,101 +1,0 @@
-library(dplyr)
-library(readr)
-library(ggplot2)
-
-eeg <- read_csv('data/preprocessed/eeg.csv')
-behavioral <- read_csv('data/preprocessed/behavioral.csv')
-
-attention.checks <- behavioral %>% 
-  filter(task=="catch", card==catch_n, phase=="test") %>%
-  select(subject_id, rt, response, task, card, phase)
-
-attention.check.responses <- attention.checks %>%
-  group_by(subject_id) %>%
-  summarize(failed = sum(is.null(rt) | (rt > 1500)))
-
-grand.averages <- eeg %>% filter(electrode %in% c("Cz", "Fz")) %>%
-  filter(good_segment==T) %>%
-  group_by(subject, t, electrode) %>%
-  summarize(m_subject = mean(v)) %>%
-  group_by(t, electrode) %>%
-  summarize(m=mean(m_subject), se=sd(m_subject)/sqrt(n()))
-
-ggplot(grand.averages, aes(x=t,y=m, ymin=m-se, ymax=m+se, color=electrode, fill=electrode))+
-  geom_ribbon(color=NA, alpha=0.1)+
-  geom_line()+
-  geom_hline(yintercept=0)+
-  geom_vline(xintercept=0)+
-  theme_minimal()+
-  scale_y_reverse()
-
-subject.eeg.counts <- eeg %>% filter(electrode %in% c("Cz", "Fz")) %>%
-  filter(good_segment==T) %>%
-  group_by(subject, hand_id, card_id) %>%
-  summarize(n=n()) %>%
-  group_by(subject) %>%
-  summarize(n=n())
-
-
-hand.data <- behavioral %>% 
-  filter(phase=='test', task=='reveal') %>%
-  mutate(subject = subject_id) %>%
-  select(-subject_id) %>%
-  group_by(subject) %>%
-  mutate(hand_id = rep(1:80, each=5)) %>%
-  select(subject, hand_id, card, card_value, wins_so_far, outcome)
-
-rpe <- hand.data %>%
-  mutate(wp_before = 1-pbinom(2-wins_so_far, 5-(card-1), 0.5), wp_after = 1-pbinom(2-(wins_so_far+card_value), 5-card, 0.5), rpe=wp_after-wp_before)
-
-entropy <- function(p){
[…]

### 2023-04-04 219-2023-analysis/manuscript.qmd
analysis with all 40 subjects!

@@
 library(ez)
+library(tidyr)
 library(osfr)
@@
     ipe == 0 ~ "None"))
+
+hand.information.minimal.set <- hand.information.with.rpe.and.ipe %>%
+  select(subject, hand_id, card_id, rpe, ipe, rpe_type, ipe_type)
 ```
@@
 
-data.eeg <- read_csv('data/preprocessed/eeg.csv')
-
-```
-
-```{r}
-#| label: Filter EEG data
-#| echo: false
-
-data.eeg.filtered <- data.eeg %>%
-  filter(good_segment == TRUE) %>%
-  filter(electrode %in% c("Cz", "Fz"))
+data.eeg.filtered <- read_csv('data/preprocessed/eeg.csv') %>%
+  filter(good_segment == TRUE, electrode %in% c("Cz", "Fz"))
 
@@
 data.eeg.filtered.rpe.ipe <- data.eeg.filtered %>%
-  left_join(hand.information.with.rpe.and.ipe, by=c("subject", "hand_id", "card_id"))
+  left_join(hand.information.minimal.set, by=c("subject", "hand_id", "card_id"))
 
@@
 
-bad.eeg.subjects <- rpe.segments %>% 
-  filter(n < 20) %>% 
-  bind_rows(ipe.segments %>% filter(n < 20)) %>%
-  pull(subject) %>%
-  unique()
+rpe.good.subjects <- rpe.segments %>%
+  group_by(subject) %>%
+  summarize(include = all(n >= 20)) %>%
+  filter(include == TRUE) %>%
+  pull(subject)
+
+ipe.good.subjects <- ipe.segments %>%
+  group_by(subject) %>%
+  summarize(include = all(n >= 20)) %>%
+  filter(include == TRUE) %>%
+  pull(subject)
 ```
@@
 We pre-registered an inclusion criteria that a participant must have at least 20 usable epochs of data of a particular event type in order to be included in the corresponding analyses.
-`r bad.eeg.subjects %>% length()` subjects were excluded for having fewer than 20 usable epochs in all event types.
-Of the remaining subjects, the average number of usable epochs for the RPE analysis in an individual subject was `r rpe.segments %>% filter(!subject %in% bad.eeg.subjects) %>% pull(n) %>% mean() %>% round(0)`, *SD*=`r rpe.segments %>% filter(!subject %in% bad.eeg.subjects) %>% pull(n) %>% sd() %>% round(0)`.
-The average number of usable epochs for the IPE analysis was `r ipe.segments %>% filter(!subject %in% bad.eeg.subjects) %>% pull(n) %>% mean() %>% round(0)`, *SD*=`r ipe.segments %>% filter(!subject %in% bad.eeg.subjects) %>% pull(n) %>% sd() %>% round(0)`.
+`r rpe.good.subjects %>% length()` subjects were included in the RPE analysis and `r ipe.good.subjects %>% length()` subjects were included in the IPE analysis.
+The average number of usable epochs for the RPE analysis in an individual subject was `r rpe.segments %>% filter(subject %in% rpe.good.subjects) %>% pull(n) %>% mean() %>% round(0)`, *SD*=`r rpe.segments %>% filter(subject %in% rpe.good.subjects) %>% pull(n) %>% sd() %>% round(0)`.
+The average number of usable epochs for the IPE analysis was `r ipe.segments %>% filter(subject %in% ipe.good.subjects) %>% pull(n) %>% mean() %>% round(0)`, *SD*=`r ipe.segments %>% filter(subject %in% ipe.good.subjects) %>% pull(n) %>% sd() %>% round(0)`.
 
@@
   filter(rpe_type != "None") %>%
-  filter(!subject %in% bad.eeg.subjects) %>%
+  filter(subject %in% rpe.good.subjects) %>%
   group_by(subject, electrode, t, rpe_type) %>%
@@
   filter(ipe_type != "None") %>%
-  filter(!subject %in% bad.eeg.subjects) %>%
+  filter(subject %in% ipe.good.subjects) %>%
   group_by(subject, electrode, t, ipe_type) %>%
@@
 ```{r}
+#| label: Difference wave summary data
+#| echo: false
+#| message: false
+
+difference.wave.data.rpe <- rpe.erp.data.subject %>%
+  pivot_wider(names_from = rpe_type, values_from = m.subject) %>%
+  mutate(difference.subject = Positive - Negative)
+
+difference.wave.summary.rpe <- difference.wave.data.rpe %>%
+  group_by(electrode, t) %>%
+  summarize(difference = mean(difference.subject), se=sd(difference.subject)/sqrt(n())) %>%
+  mutate(error_type = "Reward Prediction Error")
+
+difference.wave.data.ipe <- ipe.erp.data.subject %>%
+  pivot_wider(names_from = ipe_type, values_from = m.subject) %>%
+  mutate(difference.subject = Positive - Negative)
+
+difference.wave.summary.ipe <- difference.wave.data.ipe %>%
+  group_by(electrode, t) %>%
+  summarize(difference = mean(difference.subject), se=sd(difference.subject)/sqrt(n())) %>%
+  mutate(error_type = "Information Prediction Error")
+
+difference.erp.data <- difference.wave.summary.rpe %>% bind_rows(difference.wave.summary.ipe)
+
+```
+
+```{r}
+#| label: fig-difference-waves
+#| echo: false
+#| fig-cap: Grand average difference wave ERPs for both electrode sites (Cz, Fz).
+Values above 0 indicate that positive errors had a more positive ERP than negative errors.
+The critical time window for our analysis, 200-350ms, is highlighted.
+
+ggplot(difference.erp.data, aes(x=t,y=difference, ymin=difference-se, ymax=difference+se, color=error_type, fill=error_type))+
+  geom_hline(yintercept = 0)+
+  geom_vline(xintercept = 0)+
+  annotate("rect", xmin=min(frn.window), xmax=max(frn.window), ymin=-Inf, ymax=Inf, alpha=0.1)+
+  geom_ribbon(color=NA, alpha=0.2)+
+  geom_line()+
+   scale_color_brewer(type="qual", palette = "Set2", name="Error Type")+
+  scale_fill_brewer(type="qual", palette = "Set2", name="Error Type")+
+  facet_grid(electrode~.)+
+  labs(x="Time since card value revealed (ms)", y="Positive - Negative Prediction Errors\n(Microvolts)")+
+  theme_minimal()+
+  theme(panel.grid = element_blank())
+
+```
+
+Brydevall et al.
+reported that the ERPs in the 200-350ms window following positive prediction errors were more positive than the ERPs following negative prediction errors, for both information prediction errors and reward prediciton errors.
+Following our pre-registration, we ran two ANOVAs to see if there was a main effect of reward type (positive vs.
+negative).
+We included electrode (Cz vs.
+Fz) as a factor, as well as the interaction between reward type and electrode.
+
+```{r}
 #| label: Compute ANOVA data
@@
 
+The ANOVA table above shows that there is no significant effect of error type for reward prediction errors.
+
 ```{r}
@@
 ```
+
+However, there was a significant effect of error type for information prediction errors.
+This confirms the pattern that is evident in both @fig-grand-averages and @fig-difference-waves, where there is no clear visual difference between the ERPs for reward prediction errors, but there is a clear difference for information prediction errors.

### 2024-07-10 219-2023-analysis/manuscript.qmd
cleanup quarto headers

@@
 #| label: Load libraries
-#| echo: false
-#| message: false
-#| warning: false
+#| include: false
 
@@
 #| include: false
-#| echo: false
 
@@
 #| label: Load the Behavioral Data
-#| echo: false
-#| message: false
+#| include: false
 
@@
 #| label: Compute RPEs and IPEs
-#| echo: false
-#| message: false
+#| include: false
 
@@
 #| label: Load EEG data
-#| echo: false
-#| message: false
+#| include: false
 
@@
 #| label: Merge RPE and IPE info into EEG data
-#| echo: false
-#| message: false
+#| include: false
 
@@
 #| label: Count usable segments for RPEs
-#| echo: false
-#| message: false
+#| include: false
 
@@
 #| label: Count usable segments for IPEs
-#| echo: false
-#| message: false
+#| include: false
 
@@
 #| label: Exclude bad EEG
-#| echo: false
+#| include: false
 
@@
 #| label: Generate data for Grand Averages
-#| echo: false
-#| message: false
+#| include: false
 
@@
 #| label: Difference wave summary data
-#| echo: false
-#| message: false
+#| include: false
 
@@
 #| label: Compute ANOVA data
-#| echo: false
-#| message: false
+#| include: false
 
@@
 This confirms the pattern that is evident in both @fig-grand-averages and @fig-difference-waves, where there is no clear visual difference between the ERPs for reward prediction errors, but there is a clear difference for information prediction errors.
+
+### Exploratory Analysis
+
+Question: What if we run a model that includes both IPE and RPE as predictors to account for the fact that we have overlapping manipulations.
+And what if we include the magnitude of the prediction error rather than just presence/absence?
+
+```{r}
+#| include: false
+trial.level.data <- data.eeg.filtered.rpe.ipe %>%
+  filter(t %in% frn.window, subject %in% ipe.good.subjects) %>%
+  group_by(subject, electrode, hand_id, card_id, ipe, rpe) %>%
+  summarize(M = mean(v))
+```
+
+```{r}
+#| message: false
+library(lmerTest)
+library(brms)
+
+mm.result <- brm(M ~ rpe*electrode + ipe*electrode + (rpe + ipe + electrode | subject), data=trial.level.data)
+
+summary(mm.result)
+```
+
+```{r}
+library(broom.mixed)
+rr <- ranef(mm.result, condVar=TRUE)
+ff <- fixef(mm.result)
+aa <- broom::augment(rr)
+
+ggplot(aa,aes(estimate,level,xmin=lb,xmax=ub))+
+  geom_errorbarh(height=0)+
+  geom_vline(xintercept=0,lty=2)+
+  geom_point()+facet_wrap(~variable,scale="free_x")
+```
+
+Question: Did the reward become less salient over time? Try splitting the data based on block and look at effect of RPE on FRN.
+
+```{r}
+trial.level.data.with.block <- data.eeg.filtered.rpe.ipe %>%
+  filter(t %in% frn.window, subject %in% ipe.good.subjects) %>%
+  group_by(subject, electrode, hand_id, card_id, ipe, rpe) %>%
+  summarize(M = mean(v)) %>%
+  mutate(block = if_else(hand_id <= 40, 0, 1))
+```
+
+```{r}
+#| message: false
+library(lmerTest)
+
+mm.block.result <- lmer(M ~  rpe*block + electrode + (rpe*block + electrode | subject), data=trial.level.data.with.block)
+
+summary(mm.block.result)
+```

### 2024-07-10 219-2023-analysis/manuscript.qmd
adding trial level exploratory analysis

--- 2023-04-04 219-2023-analysis/manuscript.qmd
@@
 This confirms the pattern that is evident in both @fig-grand-averages and @fig-difference-waves, where there is no clear visual difference between the ERPs for reward prediction errors, but there is a clear difference for information prediction errors.
+
+### Exploratory Analysis
+
+```{r}
+trial.level.data <- data.eeg.filtered.rpe.ipe %>%
+  filter(subject %in% rpe.good.subjects) %>%
+  filter(t %in% frn.window) %>%
+  group_by(subject, hand_id, card_id, electrode, rpe_type, rpe, ipe) %>%
+  summarize(m.subject = mean(v))
+```
+
+```{r}
+library(brms)
+
+fit <- brm(m.subject ~ rpe + ipe + (1 + rpe + ipe | subject/hand_id), data=trial.level.data, cores=4)
+```
+
+```{r}
+summary(fit)
+```
+
+```{r}
+unsigned_fit <- brm(m.subject ~ rpe + ipe + (1 + rpe + ipe | subject/hand_id), data=trial.level.data %>% mutate(ipe=abs(ipe), rpe=abs(rpe)), cores=4)
+```
+
+```{r}
+summary(unsigned_fit)
+```
+
+```{r}
+abs.plot.data <- data.eeg.filtered.rpe.ipe %>%
+  filter(subject %in% rpe.good.subjects) %>%
+  mutate(rpe=round(rpe, 4)) %>%
+  group_by(subject, electrode, t, rpe) %>%
+  summarize(m.subject = mean(v)) %>%
+  group_by(electrode, t, rpe) %>%
+  summarize(m = mean(m.subject), se=sd(m.subject)/sqrt(n()))
+```
+
+```{r}
+ggplot(abs.plot.data, aes(x=t, y=m, ymin=m-se, ymax=m+se, color=rpe, group=rpe))+
+  geom_line()+
+  facet_wrap(~electrode)
+```
+
+```{r}
+abs.plot.data <- data.eeg.filtered.rpe.ipe %>%
+  filter(subject %in% ipe.good.subjects) %>%
+  mutate(ipe=round(ipe, 4)) %>%
+  group_by(subject, electrode, t, ipe) %>%
+  summarize(m.subject = mean(v)) %>%
+  group_by(electrode, t, ipe) %>%
+  summarize(m = mean(m.subject), se=sd(m.subject)/sqrt(n()))
+```
+
+```{r}
+ggplot(abs.plot.data, aes(x=t, y=m, ymin=m-se, ymax=m+se, color=ipe, group=ipe))+
+  geom_line()+
+  facet_wrap(~electrode)
+```
+
+```{r}
+ipe.rpe.counts <- hand.information.minimal.set %>% count(rpe, ipe)
+
+ggplot(ipe.rpe.counts, aes(x=ipe,y=rpe,size=n))+
+  geom_point()+
+  theme_minimal()
+```
+
+- Reprocess data with eye blink removal procedure
+- Use a mixed-effects model
+- Look for any RPE FRNs. (Is there literature about where else this might show up?)
+- Unsigned prediction errors?
+- Look at information gain instead of information prediction error.
+
+Descriptive statistics of RPE and IPE.
+Means and SD 
+Run the joint model, predicting FRN amplitude with RPE and IPE as predictor variables.  [We got here thinking about Camryn’s concern that we measure both in the FRN window.]
+Run first half (of the trials), second half, and random half of data to address point that the reward may have become less salient as people figured out what was going on.
+Prediction if true:  First half we detect RPE second half we don’t.
+Run a model that encodes the strength of the RPE/IPE rather than just +/-.

### 2024-07-10 219-2023-analysis/manuscript.qmd

--- 2024-07-10 219-2023-analysis/manuscript.qmd
@@
 
+```{r}
+trial.level.data <- data.eeg.filtered.rpe.ipe %>%
+  filter(subject %in% rpe.good.subjects) %>%
+  filter(t %in% frn.window) %>%
+  group_by(subject, hand_id, card_id, electrode, rpe_type, rpe, ipe) %>%
+  summarize(m.subject = mean(v))
+```
+
+```{r}
+library(brms)
+
+fit <- brm(m.subject ~ rpe + ipe + (1 + rpe + ipe | subject/hand_id), data=trial.level.data, cores=4)
+```
+
+```{r}
+summary(fit)
+```
+
+```{r}
+unsigned_fit <- brm(m.subject ~ rpe + ipe + (1 + rpe + ipe | subject/hand_id), data=trial.level.data %>% mutate(ipe=abs(ipe), rpe=abs(rpe)), cores=4)
+```
+
+```{r}
+summary(unsigned_fit)
+```
+
+```{r}
+abs.plot.data <- data.eeg.filtered.rpe.ipe %>%
+  filter(subject %in% rpe.good.subjects) %>%
+  mutate(rpe=round(rpe, 4)) %>%
+  group_by(subject, electrode, t, rpe) %>%
+  summarize(m.subject = mean(v)) %>%
+  group_by(electrode, t, rpe) %>%
+  summarize(m = mean(m.subject), se=sd(m.subject)/sqrt(n()))
+```
+
+```{r}
+ggplot(abs.plot.data, aes(x=t, y=m, ymin=m-se, ymax=m+se, color=rpe, group=rpe))+
+  geom_line()+
+  facet_wrap(~electrode)
+```
+
+```{r}
+abs.plot.data <- data.eeg.filtered.rpe.ipe %>%
+  filter(subject %in% ipe.good.subjects) %>%
+  mutate(ipe=round(ipe, 4)) %>%
+  group_by(subject, electrode, t, ipe) %>%
+  summarize(m.subject = mean(v)) %>%
+  group_by(electrode, t, ipe) %>%
+  summarize(m = mean(m.subject), se=sd(m.subject)/sqrt(n()))
+```
+
+```{r}
+ggplot(abs.plot.data, aes(x=t, y=m, ymin=m-se, ymax=m+se, color=ipe, group=ipe))+
+  geom_line()+
+  facet_wrap(~electrode)
+```
+
+```{r}
+ipe.rpe.counts <- hand.information.minimal.set %>% count(rpe, ipe)
+
+ggplot(ipe.rpe.counts, aes(x=ipe,y=rpe,size=n))+
+  geom_point()+
+  theme_minimal()
+```
+
+- Reprocess data with eye blink removal procedure
+- Use a mixed-effects model
+- Look for any RPE FRNs. (Is there literature about where else this might show up?)
+- Unsigned prediction errors?
+- Look at information gain instead of information prediction error.
+
+Descriptive statistics of RPE and IPE.
+Means and SD 
+Run the joint model, predicting FRN amplitude with RPE and IPE as predictor variables.  [We got here thinking about Camryn’s concern that we measure both in the FRN window.]
+Run first half (of the trials), second half, and random half of data to address point that the reward may have become less salient as people figured out what was going on.
+Prediction if true:  First half we detect RPE second half we don’t.
+Run a model that encodes the strength of the RPE/IPE rather than just +/-.
+
 Question: What if we run a model that includes both IPE and RPE as predictors to account for the fact that we have overlapping manipulations.
